### 2025-03-14 code/analysis-code/statistical-analysis.qmd
part 3 project v1

---
title: "Statistical analysis"
format: html
editor: visual
---

This Quarto document presents the statistical data analysis for this project, focusing on modeling. It includes bivariate analyses to identify associations between key variables and a multivariate Cox proportional hazards model to assess survival outcomes. The analysis is conducted using the tidymodels framework and survival package in R.
# Loading Package
```{r}
#load needed packages. make sure they are installed.
# Load required packages. Ensure they are installed.
library(survival)  # For Cox regression analysis
library(survminer) # For survival curve visualization
library(broom)     # For tidying model output
library(tidyverse) # For data manipulation
library(here)      # For file path management (ensures reproducibility)
library(stringr)  # Ensure string manipulation functions are available
```
# Loadinf data
```{r}
#path to data
#note the use of the here() package and not absolute paths
data_location <- here::here("data","processed-data","processeddata.rds")

#load data. 
mydata <- readRDS(data_location)
```
# Bivariate Analysis
The dataset is preprocessed to ensure consistency for survival analysis. First, all "Unknown" values in metastasis variables are removed, retaining only "Yes" and "No" categories. Next, metastasis variables are converted to factors with "No" as the reference category to ensure proper hazard ratio interpretation in Cox regression. A survival object is then created using "Survival months" as the time variable and "Vital status recode (study cutoff used)" to define event occurrence. These steps standardize the data for valid and reproducible survival analysis.

```{r}
# Remove "Unknown" from metastasis variables (keeping only "Yes" and "No")
mydata_filtered <- mydata %>%
  filter(mets_bone %in% c("Yes", "No"),
         mets_brain %in% c("Yes", "No"),
         mets_liver %in% c("Yes", "No"),
         mets_lung %in% c("Yes", "No"))

# Relevel metastasis variables so "No" is the reference category
mydata_filtered <- mydata_filtered %>%
  mutate(across(c(mets_bone, 
                  mets_brain, 
                  mets_liver, 
                  mets_lung), ~factor(.x, levels = c("No", "Yes"))))

# Create the survival object for Cox regression
surv_obj <- Surv(mydata_filtered$survival_months, mydata_filtered$vital_status == "Dead")

```

A Cox proportional hazards model is fitted separately for each metastasis variable to assess its impact on survival. The function dynamically constructs the regression formula and calculates hazard ratios (HRs) for "Yes" vs. "No" metastasis status. Results from all four models (bone, brain, liver, and lung metastases) are combined into a single table for better interpretation. The metastasis variable names are cleaned for clarity, and the final results are displayed and saved for reproducibility. This ensures a structured and systematic approach to analyzing the association between metastasis and survival outcomes.
## Bivariate Analysis: Metastasis and Survival
```{r}
####################################################
# Bivariate Analysis: Metastasis and Survival
####################################################
# Function to fit Cox regression for each metastasis variable
run_cox <- function(var) {
  formula <- as.formula(paste("surv_obj ~", paste0("`", var, "`")))  # Construct formula
  cox_model <- coxph(formula, data = mydata_filtered)  # Fit Cox model
  result <- broom::tidy(cox_model, exponentiate = TRUE)  # Convert log(HR) to HR
  return(result)  # Return model results
}

# Run Cox regression for each metastasis type (Yes vs. No comparison only)
cox_bone <- run_cox("mets_bone")
cox_brain <- run_cox("mets_brain")
cox_liver <- run_cox("mets_liver")
cox_lung <- run_cox("mets_lung")

# Combine the results into a single table
cox_results <- bind_rows(cox_bone, cox_brain, cox_liver, cox_lung)

# Rename terms for better readability
cox_results <- cox_results %>%
  mutate(term = case_when(
    term == "mets_boneYes"  ~ "Bone Metastasis",
    term == "mets_brainYes" ~ "Brain Metastasis",
    term == "mets_liverYes" ~ "Liver Metastasis",
    term == "mets_lungYes"  ~ "Lung Metastasis",
    TRUE ~ term  # Keep other terms unchanged
  ))

# Display the updated results
cox_results

# Save Cox results for reproducibility
saveRDS(cox_results, file = here::here("results", "tables", "mets_HR.rds"))

```
## KM curve Metastasis Bone
```{r}
 #Convert vital_status to a binary variable
mydata_filtered$survival_status <- ifelse(mydata_filtered$vital_status == "Dead", 1, 0)

# Create survival object
surv_obj <- Surv(mydata_filtered$survival_months, mydata_filtered$survival_status)

# Fit Kaplan-Meier survival curves for Bone Metastasis
km_fit <- survfit(surv_obj ~ mets_bone, data = mydata_filtered)

# Plot the Kaplan-Meier curves
km_plot <- ggsurvplot(
  km_fit,
  data = mydata_filtered,
  conf.int = TRUE,        # Show confidence intervals
  pval = TRUE,            # Display p-value
  risk.table = FALSE,      # Show risk table below the plot
  legend.title = "Bone Metastasis",
  legend.labs = c("No Bone Metastasis", "Bone Metastasis"),
  xlab = "Time (months)", # Adjust based on your dataset
  ylab = "Survival Probability",
  ggtheme = theme_minimal(),

)

# Print the KM plot
print(km_plot)

# Save the plot
ggsave(filename = here::here("results", "figures", "KM_curve_bone_mets.png"), plot = km_plot$plot, dpi = 300, width = 8, height = 6)

```
## KM curve Metastasis Brain
```{r}
 ####################################################
# Kaplan-Meier Analysis: Brain Metastasis and Survival
####################################################

# Load required libraries
library(survival)
library(survminer)

# Convert vital_status to a binary variable
mydata_filtered$survival_status <- ifelse(mydata_filtered$vital_status == "Dead", 1, 0)

# Create survival object
surv_obj <- Surv(mydata_filtered$survival_months, mydata_filtered$survival_status)

# Fit Kaplan-Meier survival curves for Brain Metastasis
km_fit <- survfit(surv_obj ~ mets_brain, data = mydata_filtered)

# Plot the Kaplan-Meier curves
km_plot <- ggsurvplot(
  km_fit,
  data = mydata_filtered,
  conf.int = TRUE,        # Show confidence intervals
  pval = TRUE,            # Display p-value
  risk.table = FALSE,     # Hide risk table (Set to TRUE if needed)
  legend.title = "Brain Metastasis",
  legend.labs = c("No Brain Metastasis", "Brain Metastasis"),
  xlab = "Time (months)", # Adjust based on your dataset
  ylab = "Survival Probability",
  ggtheme = theme_minimal()
)

# Print the KM plot
print(km_plot)

# Save the plot
ggsave(filename = here::here("results", "figures", "KM_curve_brain_mets.png"), plot = km_plot$plot, dpi = 300, width = 8, height = 6)

```
## KM curve Metastasis Liver
```{r}
####################################################
# Kaplan-Meier Analysis: Liver Metastasis and Survival
####################################################

# Load required libraries
library(survival)
library(survminer)

# Convert vital_status to a binary variable
mydata_filtered$survival_status <- ifelse(mydata_filtered$vital_status == "Dead", 1, 0)

# Create survival object
surv_obj <- Surv(mydata_filtered$survival_months, mydata_filtered$survival_status)

# Fit Kaplan-Meier survival curves for Liver Metastasis
km_fit <- survfit(surv_obj ~ mets_liver, data = mydata_filtered)

# Plot the Kaplan-Meier curves
km_plot <- ggsurvplot(
  km_fit,
  data = mydata_filtered,
  conf.int = TRUE,        # Show confidence intervals
  pval = TRUE,            # Display p-value
  risk.table = FALSE,     # Hide risk table (Set to TRUE if needed)
  legend.title = "Liver Metastasis",
  legend.labs = c("No Liver Metastasis", "Liver Metastasis"),
  xlab = "Time (months)", # Adjust based on your dataset
  ylab = "Survival Probability",
  ggtheme = theme_minimal()
)

# Print the KM plot
print(km_plot)

# Save the plot
ggsave(filename = here::here("results", "figures", "KM_curve_liver_mets.png"), plot = km_plot$plot, dpi = 300, width = 8, height = 6)

```

## KM curve Metastasis Lung 
```{r}
####################################################
# Kaplan-Meier Analysis: Lung Metastasis and Survival
####################################################

# Load required libraries
library(survival)
library(survminer)

# Convert vital_status to a binary variable
mydata_filtered$survival_status <- ifelse(mydata_filtered$vital_status == "Dead", 1, 0)

# Create survival object
surv_obj <- Surv(mydata_filtered$survival_months, mydata_filtered$survival_status)

# Fit Kaplan-Meier survival curves for Lung Metastasis
km_fit <- survfit(surv_obj ~ mets_lung, data = mydata_filtered)

# Plot the Kaplan-Meier curves
km_plot <- ggsurvplot(
  km_fit,
  data = mydata_filtered,
  conf.int = TRUE,        # Show confidence intervals
  pval = TRUE,            # Display p-value
  risk.table = FALSE,     # Hide risk table (Set to TRUE if needed)
  legend.title = "Lung Metastasis",
  legend.labs = c("No Lung Metastasis", "Lung Metastasis"),
  xlab = "Time (months)", # Adjust based on your dataset
  ylab = "Survival Probability",
  ggtheme = theme_minimal()
)

# Print the KM plot
print(km_plot)

# Save the plot
ggsave(filename = here::here("results", "figures", "KM_curve_lung_mets.png"), plot = km_plot$plot, dpi = 300, width = 8, height = 6)
```

## Bivariate Analysis: Race and Survival

The race variable is recoded as a factor with "White" as the reference category to ensure proper comparison in Cox regression. A separate model is fitted to assess the association between race and survival, with hazard ratios (HRs) calculated for Black, Asian or Pacific Islander, and American Indian/Alaska Native populations relative to White. The results are formatted for clarity, renaming the terms to indicate direct comparisons. Finally, the updated results are displayed and saved for reproducibility, allowing for structured and interpretable survival analysis by race.
```{r}
####################################################
# Bivariate Analysis: Race and Survival
####################################################

# Recode the race variable as a factor, ensuring "White" is the reference category
mydata_filtered <- mydata_filtered %>%
  mutate(race = factor(race, 
                                                 levels = c("White", "Black", "Asian or Pacific Islander", "American Indian/Alaska Native")))

# Function to fit Cox regression for Race categories
run_cox_race <- function() {
  # Construct Cox model formula using race variable
  formula <- as.formula("surv_obj ~ race")

  # Fit Cox model
  cox_model <- coxph(formula, data = mydata_filtered)

  # Convert output to tidy format and exponentiate coefficients to get Hazard Ratios (HRs)
  result <- broom::tidy(cox_model, exponentiate = TRUE)

  return(result)  # Return model results
}

# Run the Cox regression for Race categories
cox_race_results <- run_cox_race()

# Rename terms for better readability
cox_race_results <- cox_race_results %>%
  mutate(term = case_when(
    term == "raceBlack" ~ "Black vs. White",
    term == "raceAsian or Pacific Islander" ~ "Asian vs. White",
    term == "raceAmerican Indian/Alaska Native" ~ "American Indian vs. White",
    TRUE ~ term  # Keep other terms unchanged
  ))

# Display results
cox_race_results

# Save the results for reproducibility
saveRDS(cox_race_results, file = here::here("results", "tables", "race_HR.rds"))

```

## Kaplan-Meier Survival Curve: Race and Survival

```{r}
####################################################
# Kaplan-Meier Survival Curve: Race and Survival
####################################################

# Ensure survival variables are properly formatted
mydata_filtered <- mydata_filtered %>% 
  mutate(survival_months = as.numeric(survival_months),
         vital_status = ifelse(vital_status == "Dead", 1, 0))

# Handle NA values in survival_months
if(any(is.na(mydata_filtered$survival_months))) {
  cat("Warning: NAs introduced in survival_months due to coercion. Removing NAs for survival analysis.\n")
  mydata_filtered <- mydata_filtered %>% filter(!is.na(survival_months))
}

# Define survival object
surv_obj <- Surv(mydata_filtered$survival_months, mydata_filtered$vital_status)

# Fit Kaplan-Meier survival model for Race
fit_race <- survfit(surv_obj ~ race, data = mydata_filtered)

# Generate Kaplan-Meier plot for Race categories
ggsurvplot(fit_race, data = mydata_filtered, conf.int = TRUE, pval = TRUE, risk.table = TRUE, 
           title = "Kaplan-Meier Survival Curves by Race", 
           xlab = "Time in Months", ylab = "Survival Probability",
           legend.title = "Race",
           legend.labs = levels(mydata_filtered$race), # Dynamically assign race labels
           xlim = c(0, 130), ggtheme = theme_minimal())
# Save the plot
ggsave(filename = here::here("results", "figures", "KM_curve_race.png"), plot = km_plot$plot, dpi = 300, width = 8, height = 6)
```

## Bivariate Analysis: Age and Survival
The age variable is recoded as a factor with "20-24 years" set as the reference category to allow for meaningful comparisons in Cox regression. A separate model is fitted to evaluate the association between different age groups and survival outcomes, with hazard ratios (HRs) estimated relative to the 20-24 years category. The results are formatted to improve clarity, renaming each term to reflect direct age group comparisons. Finally, the updated results are displayed and saved for reproducibility, ensuring a structured and interpretable analysis of age-related survival differences.

```{r}
####################################################
# Bivariate Analysis: Age and Survival
####################################################

# Function to fit Cox regression for Age groups
run_cox_age <- function() {
  # Construct Cox model formula using the updated age variable name
  formula <- as.formula("surv_obj ~ age_group")

  # Fit Cox model
  cox_model <- coxph(formula, data = mydata_filtered)

  # Convert output to tidy format and exponentiate coefficients to get Hazard Ratios (HRs)
  result <- broom::tidy(cox_model, exponentiate = TRUE)

  return(result)  # Return model results
}

# Run the Cox regression for Age categories
cox_age_results <- run_cox_age()

# Rename terms for better readability (comparing each group to 20-24 years)
cox_age_results <- cox_age_results %>%
  mutate(term = str_replace_all(term, "`", "")) %>%
  mutate(term = case_when(
    str_detect(term, "age_group25-29 years") ~ "25-29",
    str_detect(term, "age_group30-34 years") ~ "30-34",
    str_detect(term, "age_group35-39 years") ~ "35-39",
    str_detect(term, "age_group40-44 years") ~ "40-44",
    str_detect(term, "age_group45-49 years") ~ "45-49",
    str_detect(term, "age_group50-54 years") ~ "50-54",
    str_detect(term, "age_group55-59 years") ~ "55-59",
    str_detect(term, "age_group60-64 years") ~ "60-64",
    str_detect(term, "age_group65-69 years") ~ "65-69",
    str_detect(term, "age_group70-74 years") ~ "70-74",
    str_detect(term, "age_group75-79 years") ~ "75-79",
    str_detect(term, "age_group80-84 years") ~ "80-84",
    str_detect(term, "age_group85+ years") ~ "85+",
    TRUE ~ term  # Keep other terms unchanged
  ))

# Display results
cox_age_results

# Save the results for reproducibility
saveRDS(cox_age_results, file = here::here("results", "tables", "age_HR.rds"))

```
## Bivariate Analysis: Year of Diagnosis and Survival

```{r}
####################################################
# Bivariate Analysis: Year of Diagnosis and Survival
####################################################

# Identify the earliest year available in the dataset to set as reference
reference_year <- min(as.numeric(mydata_filtered$diagnosis_year), na.rm = TRUE)

# Convert Year of Diagnosis to a factor with the reference year as baseline
mydata_filtered <- mydata_filtered %>%
  mutate(diagnosis_year = factor(diagnosis_year, 
                                 levels = sort(unique(diagnosis_year), decreasing = FALSE)))

# Function to fit Cox regression for Year of Diagnosis
run_cox_year <- function() {
  # Construct Cox model formula using diagnosis_year variable
  formula <- as.formula("surv_obj ~ diagnosis_year")

  # Fit Cox model
  cox_model <- coxph(formula, data = mydata_filtered)

  # Convert output to tidy format and exponentiate coefficients to get Hazard Ratios (HRs)
  result <- broom::tidy(cox_model, exponentiate = TRUE)

  return(result)  # Return model results
}

# Run the Cox regression for Year of Diagnosis
cox_year_results <- run_cox_year()

# Rename terms for better readability (comparing each year to the reference year)
cox_year_results <- cox_year_results %>%
  mutate(term = str_replace_all(term, "`", "")) %>%
  mutate(term = case_when(
    TRUE ~ paste0(term, " vs. ", reference_year)  # Compare each year with the reference
  ))

# Display results
cox_year_results

# Save the results for reproducibility
saveRDS(cox_year_results, file = here::here("results", "tables", "year_HR.rds"))

```

## Bivariate Analysis: Median Household Income (Inflation-Adjusted to 2019)

The median household income variable, inflation-adjusted to 2019, is recoded as a factor with "$75,000+" as the reference category to allow for clear comparisons in Cox regression. A separate model is fitted to assess the association between income levels and survival outcomes, with hazard ratios (HRs) estimated relative to the highest income group. The results are formatted for better readability by appending comparison labels for each income group against "$75,000+." Finally, the updated results are displayed and saved for reproducibility, ensuring a structured and interpretable analysis of socioeconomic disparities in survival.

```{r}
####################################################
# Bivariate Analysis: Median Household Income (Inflation-Adjusted to 2019)
####################################################
# Convert Household Income to a factor with "$75,000+" as the reference category
mydata_filtered <- mydata_filtered %>%
  mutate(median_income = factor(median_income, 
                                                                  levels = c("$75,000+", 
                                                                             "$35,000 - $39,999", 
                                                                             "$40,000 - $44,999", 
                                                                             "$45,000 - $49,999", 
                                                                             "$50,000 - $54,999",
                                                                             "$55,000 - $59,999",
                                                                             "$60,000 - $64,999",
                                                                             "$65,000 - $69,999",
                                                                             "$70,000 - $74,999")))

# Function to fit Cox regression for Household Income
run_cox_income <- function() {
  # Construct Cox model formula using Household Income variable
  formula <- as.formula("surv_obj ~ median_income")

  # Fit Cox model
  cox_model <- coxph(formula, data = mydata_filtered)

  # Convert output to tidy format and exponentiate coefficients to get Hazard Ratios (HRs)
  result <- broom::tidy(cox_model, exponentiate = TRUE)

  # Add column for variable reference
  result$Variable <- "Household Income"

  return(result)  # Return model results
}

# Run the Cox regression for Household Income
cox_income_results <- run_cox_income()

# Rename terms for better readability (comparing each income group to "$75,000+")
cox_income_results <- cox_income_results %>%
  mutate(term = str_replace_all(term, "`", "")) %>%
  mutate(term = case_when(
    TRUE ~ paste0(term, " vs. $75,000+")
  ))

# Display results
cox_income_results

# Save the results for reproducibility
saveRDS(cox_income_results, file = here::here("results", "tables", "income_HR_75K.rds"))

```

# Multivariable Cox Proportional Hazards Model
```{r}
####################################################
# Multivariable Cox Proportional Hazards Model
####################################################

# Define the multivariable Cox regression formula with all selected covariates
multivariable_formula <- as.formula("surv_obj ~ 
  mets_bone + 
  mets_brain + 
  mets_liver + 
  mets_lung + 
  race + 
  age_group + 
  diagnosis_year + 
  median_income")

# Fit the Cox proportional hazards model
multivariable_cox <- coxph(multivariable_formula, data = mydata_filtered)

# Convert model output to a tidy format and exponentiate coefficients to get Hazard Ratios (HRs)
multivariable_results <- broom::tidy(multivariable_cox, exponentiate = TRUE)

# Rename terms for better readability in the output
multivariable_results <- multivariable_results %>%
  mutate(term = str_replace_all(term, "`", "")) %>%
  mutate(term = case_when(
    term == "mets_boneYes" ~ "Bone Metastasis",
    term == "mets_brainYes" ~ "Brain Metastasis",
    term == "mets_liverYes" ~ "Liver Metastasis",
    term == "mets_lungYes" ~ "Lung Metastasis",
    term == "raceBlack" ~ "Black",
    term == "raceAsian or Pacific Islander" ~ "Asian or Pacific Islander",
    term == "raceAmerican Indian/Alaska Native" ~ "American Indian/Alaska Native",
    str_detect(term, "age_group") ~ str_remove(term, "age_group") %>% str_remove(" vs. 20-24 years"),
    str_detect(term, "household_income") ~ str_remove(term, "household_income") %>% str_remove(" vs. \\$75,000\\+"),
    TRUE ~ term  # Keep other terms unchanged
  ))

# Display the multivariable Cox model results
multivariable_results

# Save the results for reproducibility
saveRDS(multivariable_results, file = here::here("results", "tables", "multivariable_cox_HR.rds"))

```