### 2025-04-26 reviews/zuber-review-byhopeg.qmd

---
title: Project Review Template 
author: Hope Grismer
date: date-modified
format: 
  html:
    toc: true
    toc-depth: 3
    number-sections: true
---

# Overview

Title of project: The Clinicopathological Features and Survival Outcomes of Patients with Different Metastatic Sites in Stage IV Endometrial Cancer: A SEER-Based Cohort Study

Name of project author(s): Mohammed Zuber

Name of project reviewer: Hope Grismer

# Instructions

Write your comments and feedback below for each section/component of the project. The goal should be to help the author improve their project. Make comments as constructive and actionable as possible. You can provide both criticism and praise.

For each component, pick one summary statement by deleting the ones that do not apply and keeping only the one that you think most closely summarizes a given component.

Make sure your final document compiles/renders into a readable, well-formatted html document.

Delete any sections/text of this template that are not part of your final review document. (Including these instructions.)

# Specific project content evaluation

## Background, Context and Motivation

How well is the context of the project described? Is a comprehensive background, including summary of previous/related work given? Is the project well placed into the context of existing work (including proper referencing of existing work). Is it clear why the project was undertaken and what new information it hopes to provide?

### Feedback and Comments

The manuscript provides strong background information, including clear statistics about incidence and survival rates of endometrial cancer. Prior studies are referenced appropriately. The motivation for the study — understanding survival outcomes by metastasis site — is well established and timely. The importance of investigating metastatic patterns is clearly justified given the survival disparities highlighted.

Minor note: A little more explicit tie-in about what gaps remain in the current literature would make the motivation even stronger (e.g., “Previous studies have not systematically compared brain, liver, bone, and lung outcomes in a population-based cohort”).

### Summary assessment (PICK ONE, DELETE THE OTHERS)

-   strong contextualization and motivation

## Question description

How well and clear are the question(s)/hypotheses the project aims to address described? Is it clear how the questions relate to the data?

### Feedback and Comments

The research question is clearly stated and fully aligned with the study goals. Both the question ("What are the survival outcomes across metastatic sites?") and the hypothesis ("different sites have different survival implications") are logically connected to the data being analyzed. Great job stating not only the expected differences but how they might arise (tumor behavior, treatment responses).

### Summary assessment

-   question/hypotheses fully clear

## Data description

How well is the data overall described? Is the source provided? Is a codebook or other meta-information available that makes it clear what the data is?

### Feedback and Comments

The SEER database isv**well described**, including when it started capturing metastasis-specific data (2010), and the exact fields pulled for analysis.

Minor nitpick: There’s a typo where "brain metastasis" is listed twice in the extracted variables section.

### Summary assessment

-   source and overall structure of data well explained

## Data wrangling and exploratory analysis

How well is the data cleaned/processed and explored? Are all steps reasonable and well explained? Are alternatives discussed and considered? Are meaningful exploratory results shown (e.g. in the supplementary materials)?

### Feedback and Comments

The cleaning steps are reasonably explained:

-   Exclusions (age \<18, non-endometrial deaths, early-stage cancers) are sensible

-   Variable formatting (categorical to factors) is described

-   Missingness handling is addressed

The exploratory analysis includes descriptive stats, tables, and good plots for survival time, metastasis site, and demographics.

Minor room for improvement:

-   Could use slightly more commenting in the code (especially in processingcodemz)

-   Exploratory analysis focuses mainly on survival; a few more EDA plots (e.g., cross-tabulation of metastasis by race) could deepen understanding.

### Summary assessment

-   some weaknesses in wrangling and exploratory component

## Appropriateness of Analysis

Were the analysis methods appropriate for the data? Was the analysis done properly? Were different components of the analysis (e.g. performance measure, variable selection, data pre-processing, model evaluation) done in the best way possible and explained well?

### Feedback and Comments

The statistical analysis is appropriate and well implemented:

-   Kaplan-Meier curves are correctly used for unadjusted survival comparisons

-   Bivariate and multivariable Cox models are used correctly

-   Random Survival Forest model is a strong (and justified) complement to traditional modeling

Model performance (AUC = 0.74) is respectable. Liked your choice to focus on cause-specific survival rather than all-cause mortality.

### Summary assessment

-   strong and reasonable analysis

## Presentation

How well are results presented? Are tables and figures easy to read and understand? Are the main figures/tables publication level quality?

### Feedback and Comments

Tables and figures are clear and logically described in the text, even though the actual figures were not attached to the Word document.\
(From reading the code, I can tell they would render well — e.g., ggplot2 survival curves, Random Forest variable importance plots.)

A bit more consistent labeling of axes/titles would elevate figures to true publication quality.

### Summary assessment

-   results are very well presented (tiny room for improvement)

## Discussion/Conclusions

Are the study findings properly discussed? Are strengths and limitations acknowledged? Are findings interpreted properly?

### Feedback and Comments

The discussion is thorough and well contextualized:

-   Main findings (brain/liver worse prognosis) are clearly summarized

-   Racial and age disparities are thoughtfully discussed

-   Strengths (population-based sample) and limitations (no treatment data) are acknowledged

The interpretation is responsible — no overstatement of the findings. References are integrated nicely.

### Summary assessment

-   strong, complete and clear discussion

## Further comments

*Add any other comments regarding the different aspects of the project here. Write anything you think can help your classmate improve their project.*

# Overall project content evaluation

Evaluate overall features of the project by filling in the sections below.

## Structure

Is the project well structured? Are files in well labeled folders? Do files have reasonable names? Are all "junk" files not needed for analysis/reproduction removed? By just looking at files and folders, can you get an idea of how things fit together?

### Feedback and Comments

-   Three main Quarto files corresponding to cleaning, EDA, and analysis

-   No junk files

-   Clear naming conventions

-   Could organize some things into subfiles and folders???

### Summary assessment

-   Well structured

## Documentation

How well is the project documented? Are you able to understand each step of the whole analysis, each decision that was made, and each line of code? Is enough information provided as comments in code or as part of Rmd files?

### Feedback and Comments

The manuscript is extremely well documented.\
The code is moderately documented

### Summary assessment

-   decently documented with some gaps

## Reproducibility

Are all results fully reproducible? Is documentation provided which clearly explains how to reproduce things, and does it work without the need for any manual intervention? Are you able to re-run the whole analysis without having to do manual interventions/edits?

### Feedback and Comments

-   All code files are provided

-   Data source (SEER) is identified

-   GitHub repo link is included

### Summary assessment

-   fully reproducible without issues

## Thoroughness

How thorough was the overall study? Were alternatives (e.g. different ways of processing the data or different models) considered? Were alternatives discussed? Were the questions/hypotheses fully and thoroughly addressed?

### Feedback and Comments

-   Both traditional statistics and machine learning were applied

-   Multiple predictors were considered

-   Sensitivity to race, age, and income differences were discussed

-   Strengths and limitations were there too

### Summary assessment

-   strong level of thorougness

## Further comments

Thank you for an amazing project! I really was inspired in getting to look yours over. It was super well-rounded project– very close to publication-level. I especially liked the depth and strength of writing in your manuscript.